\documentclass[
  12pt,
  oneside,
  a4paper
  ]{report}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\usepackage{inputenc}
\usepackage[
  rmargin=2.5cm,
  lmargin=3.5cm,
  tmargin=2.5cm,
  bmargin=2.5cm
  ]{geometry}
% \usepackage{showframe}
\usepackage{thesis}

%%%%%%%%%%%%%%%%%%%%%%%%
% Thesis Main Document %
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{}
\author{Ali Osman Berk Şapcı}
\maketitle

\tableofcontents
\listoffigures
\listoftables

\begin{abstract}
	Sleep is an essential behavioral program conserved across the animal kingdom.
	In order to understand the underlying function of sleep, careful characterization of the changes in behavior and physiology is needed in powerful genetic model systems such as \textit{Drosophila Melanogaster}.
	Recent advances in machine learning have enabled tracking of the body parts and robust pose estimation; however, automated quantification of the behaviors requires providing a mapping from a pair of spatial coordinates to behavioral categories.
	Existing methods and pipelines are developed with behaviors defined by major postural changes in mind.
	The task of  detecting and successfully mapping behaviors exhibited during sleep comes with its unique challenges.
	Our task of phenotyping sleep requires tackling with behaviors defined by unobtrusive changes, and that sparsely occur during long sleep cycles.
	To this end, we develop \texttt{basty} (Automated \underline{B}ehavioral \underline{A}nalysis of A\underline{s}leep Frui\underline{t} Fl\underline{y}); a novel, end-to-end pipeline, which is made public as a configurable, open source and easy-to-use software package. Our pipeline consists of multiple stages; such as computing meaningful behavioral representations, detecting activities in long sleep experiments (14-16 hours), and nearest neighbors analysis in a low dimensional behavioral space. We evaluate our pipeline with a dataset consisting of sleep-deprivation and wild-type sleep experiments; considering five behavioral categories. Results show that our pipeline successfully maps behavioral categories, achieving an AUC score of 0.8, and it is able to detect unseen behaviors and differences in behavioral repertoires.
	Furthermore, we present a brief analysis of the behavioral repertoire exhibited during sleep by examining spatio-temporal characteristics of the behaviors and their temporal organization.
	% A \textit{Bastı} (English: \textit{Basty}, Azerbaijani Turkish: \textit{Basdı}, Anatolian Turkish: \textit{Basdırık}) is an evil spirit in Turkic mythology which rides on people's chests while they sleep, bringing on  nightmares.
	% Bastı sits astride a sleeper's chest and becomes heavier until the crushing weight awakens the terrified and breathless dreamer.
	% The victim awakes, unable to move under the Bastıs weight.
	% It may also include lucid dreams.
	% There are different types of Bastı in Anatolia: Al-Bastı, Kara-Bastı, Kul-Bastı, Sarı-Bastı.
\end{abstract}

\input{1-Introduction}
\input{2-DataCollection}
\input{3-FeatureExtraction}
\input{4-ActivityDetection}
\input{5-BehaviorMapping}
\input{6-Results}
\input{7-BastySoftwarePackage}
\input{8-Conclusion}

\bibliography{refs}

\end{document}
