\setlength{\parindent}{0pt}
\chapter{\bf CONCLUSION}\label{chapter:conclusion}

% Retell what you have done.
Ethology, the scientific study of animal behavior, focuses on behavior under natural conditions and is a rapidly growing field.
With its scientific roots in the late 19th and 20th centuries, it became a well-recognized scientific discipline and yielded significant insight into general principles of the nervous system.
Behavior, as arguably the most robust output of the brain, is needed to be deciphered to understand complex neurobiological phenomena, including sleep.
Technical advances, ranging from imaging technologies to machine learning techniques, transformed ethology into a new computational discipline, and the field of computational ethology has emerged.

In the past decade, remarkable tools, which involve computer vision and machine learning, have been developed for automated quantification of animal behavior.
Employing such tools provided unprecedented insight into complex behaviors, such as hunting and mating \citep{mearns_deconstructing_2020, janisch_deciphering_2021}.
Yet still, challenges remain open.
While behaviors exhibited by awake animals tend to be characterized by major postural and positional changes, the behavioral repertoire of asleep animals includes subtle movements and unobtrusive changes that sparsely occur during long sleep cycles.
Thus, exploring and quantifying the behavioral repertoire of sleep requires special care and additional considerations.

To this end, we develop a multi-stage and end-to-end behavior mapping pipeline, specifically designed for animals in dormancy state, and release it as a configurable and easy-to-use software package, called \texttt{basty}.
Our approach consists of computing representations capturing complex spatio-temporal characteristics of behaviors, detecting micro-activities exhibited in long-duration sleep bouts, and mapping detected activities utilizing a nearest neighbor based analysis on the low-dimensional behavioral embeddings.
Our pipeline is able to handle noisy pose estimation data, with missing points (i.e., occlusions), through rigorous pre-processing and post-processing steps.
Using novel approaches such as semi-supervised pair embeddings approach and experiment committee by voting, \texttt{basty} can tackle mapping of the rich behavioral repertoire of dormant fruit fly with limited supervision.

We evaluate our \texttt{basty} with a dataset of sleep deprivation and wild-type sleep experiments, focusing on five behavioral categories: grooming, feeding, proboscis pumping, haltere switch, and postural adjustments.
Automated quantification of proboscis pumping and haltere switch is achieved for the first time in this work.
Such behaviors are potentially related to behaviorally defined states of sleep similar to the REM sleep (rapid eye movement sleep) phase observed in mammals and birds.
Results demonstrate that our pipeline robustly detects activities, and successfully maps behavioral categories.
Overall, \texttt{basty} achieves a macro average of 0.8 AUC score.
Moreover, we demonstrate its ability to detect unobserved behaviors and differences between behavioral repertoires.

Although we focus on 5 behavioral categories, in fact, the behavioral repertoire of dormant Drosophila Melanogaster is much richer. There exist behaviors that do not fall into any of 5 the defined categories. 
As behaviors need to be annotated, the problem of ``lumping'' versus ``splitting'' emerges due to the nature of behaviors. For instance, we lump face grooming and abdomen grooming under a single category of grooming.
Therefore, a future direction is extending the considered behavioral categories, and even splitting existing categories into more strictly defined subcategories. 

The approach of segmentation of continuous and hierarchical behaviors into categorical components is necessarily human-biased and does not reflect the true nature of behaviors. A sequence of simple behaviors together might hierarchically generate compositional behaviors. Similarly, two behaviors might accompany each other, resulting in a composite behavior. One can utilize behavioral scores to analyze such structures, however, in its current state, our pipeline does not directly deal with composite and compositional behaviors.
Hence, \texttt{basty} can be extended to analyze such behaviors in the future.

Popular pose estimation tools, such as DeepLabCut and SLEAP, perform tracking on the videos recorded with a single camera, and hence, the inference is done in the two-dimensional projection. The resulting pose estimation data does directly reflect the true spatial characteristics and contains  many missing points due to occlusions. However, the true spatial characteristics of behaviors can only be captured in three-dimensional space.
Our pipeline, \texttt{basty}, can easily be configured to operate on three-dimensional pose values, as three-dimensional pose estimation tools become more available in the future.

% How does this research enables other research, sleep phenotyping, molecular biology experiments to identify genetic programs, etc.