\clearpage
\setlength{\parindent}{0pt}
\begin{center}
	\MakeUppercase{\textbf{acknowledgment}} \\ [3\baselineskip]
\end{center}

% Öznur Taştan
I honestly couldn't ask for a better advisor than Öznur Taştan.
Being her student and working closely with her was a privilege.
I am grateful for her invaluable guidance, generous support, and sincerity.
Her impact is immeasurable, and it is a blessing to have an exceptional advisor like her.
Your role in my life has been more than an academic advisor.
Değerli hocam, thank you for believing in me.

% Sündüz Keleş
I would like to express reverence and my gratitude to my co-advisor, Sündüz Keleş, for being an excellent role model and an incredible mentor.
I will be forever indebted to her, whose enthusiasm and determinism have always been motivating, and who has inspired me beyond my career and education.
Thank you for encouraging me, and providing precious opportunities.

% Mehmet Fatih Keleş
Danke.

I would like to thank Onur Varol, Ercüment Çiçek, and Gözde Bozdağı Akar, for the time they spent while serving on my committee.
Their valuable feedback and critical evaluations have improved the quality and the readability of this work.

% Enes Recep Türkoğlu
% Thanks to my dear friend and roommate Enes R. Türkoğlu for being a constant but fun source of procrastination.
% It has been seven good years; time does not pass, it continues.

% Family
I couldn't have achieved anything without my family.
I cannot express how grateful I am for having the most supporting and lovely parents and sister.
They have raised me as an individual who aims to contribute to greater things and who does not give up easily.
My father, Recai, was there for me in both good times and bad times.
He worked very hard to make sure that I had the best education possible.
The continuous and unconditional love of my mother, Fatma Hanım, has supported me at every stage of my life, and helped me to ``stay in line''.
My little sister, Elif Nehir, has always cheered me up and her existence itself is a source of peace and joy.

% Selma
Thanks.
